---
title: "CAPITULO 7"
output: rmarkdown::github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####loading packages
```{r}
library(tidyverse)
library(corrplot)
library(modelr)
```

#### categorical variable

```{r}
diamonds %>% ggplot() + geom_bar(mapping = aes(x = cut))
```

#### A altura de cada coluna representa quantas observações ocorreram em cada valor, é possível contar esses valores com count()

```{r}
diamonds %>% count(cut)
```

#### Variável Contínua
#### São variáveis que podem conter qualquer tipo de valores infinitos. Números e datas são exemplos de variáveis continuas.

#### Para avaliar distribuição utiliza-se histograma

```{r}
diamonds %>% ggplot() + geom_histogram(mapping = aes(price), bins = 15)
```

#### em nosso dataset as variáveis contínuas são: Carat,depth,table,price,x,y,z
```{r}
head(diamonds)
```

```{r}
diamonds %>% 
  count(cut_width(carat, 0.5))
```
#### O geom_freqpoly pode use ser utilizado como histograma.


```{r}
diamonds %>% ggplot() + geom_histogram(mapping = aes(x = price, fill = cut))
```

```{r}
diamonds %>% ggplot() + geom_freqpoly(mapping = aes(x = carat, color = cut))
```

```{r}
names(diamonds)
diamonds %>% ggplot() + geom_boxplot(mapping = aes(x = cut, y = price, color = cut))
```

## 7.3.4
#### Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? #### Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
dimensions = diamonds[,c('x','y','z')]
stacked = stack(dimensions)
stacked$ind = as.factor(stacked$ind)
stacked %>% ggplot() + geom_density(mapping = aes(x = values, fill = ind)) + facet_wrap(~ind)
```
### Como podemos ver o X está mais distribuido e o Z mais concentrado, provavelmente o X apresenta um maior desvio padrão do que Y, podemos verificar esta hipotse com a tabela abaixo.

```{r}
by(stacked$values, stacked$ind, sd)
```
### Conforme pode ser visto o desvio padrão de X é maior do que de Z. 


#### Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
diamonds
diamonds %>% ggplot() + geom_histogram(mapping = aes(x = price),binwidth = 1000, color = 'black', fill = 'grey')
```

#### Conforme pode ser visto os dados não apresentam distribuição normal, estão mais concentrados em valores menores (positive skeweded).
#### podemos fazer o teste de kolmogorov smirnov

```{r}
diamonds$price = as.numeric(diamonds$price)
diamoprice = diamonds[1:4999,]

testekol = shapiro.test(diamoprice$price)
testekol
#valor menor do que 0.05 o que confirma a hipotse que os dados não estão distribuidos normal.

```

### how many diamonds are 0.99 carat, how many are 1 carat?

```{r}
names(diamonds)
diamonds %>% group_by(carat) %>% filter(carat == 0.99 | carat == 1) %>% summarise(n = n()) 
```

#### removing NA (Not recommended this way)
```{r}
diamonds2 = diamonds %>% filter(between(y,3,20))
diamonds2

```

####Removing NA (recommended way)

```{r}
diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
colSums(is.na(diamonds2))
```
#### 9 values out of the bounds (<3 or >20)

```{r}
diamonds2 %>% ggplot() + geom_point(mapping = aes(x = x, y = y))
```
#### still veriyinf some outliers
```{r}

nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  ) %>% 
  ggplot(mapping = aes(sched_dep_time)) + 
    geom_freqpoly(mapping = aes(colour = cancelled), binwidth = 1/4)

```
##### 7.4.1 Exercises 

```{r}
diamonds2 %>% ggplot() + geom_histogram(mapping = aes(x = y))
```
##### we can see that when there's NA value there a space between bins.

### If variation describes the behavior within a variable, covariation describes the behavior between variables.

```{r}
ggplot(data = diamonds, aes(x = price, y = ..density..)) + geom_freqpoly(mapping = aes(color = cut))
```

#### another way of doing it
```{r}
diamonds %>% ggplot() + geom_density(mapping = aes(x = price, fill = cut), alpha = 0.4)
```

### data of the chart above are stardardized so its possible to see the differences, otherwise it wouldn't be that easy. 

```{r}
diamonds2 %>%  ggplot() + geom_boxplot(mapping = aes(x = cut,y= price))
```

### What variable in the diamonds dataset is most important for predicting the price of a diamond?
### How is that variable correlated with cut? Why does the combination of those two relationships ### lead to lower quality diamonds being more expensive?
```{r}
diamonds3 = diamonds[,c('carat','depth','table','price','x','y','z')]
corrplot(cor(diamonds3), method  = 'number')

```
#### as it's showed w/ price the variables which correlate the most are carat, x,y,z, for predicting we could use carat, x,y,z

install.packages('polycor')
#### as 

```{r}
diamonds
diamonds %>% ggplot() + geom_boxplot(mapping = aes(x = cut, y = price))
```
#### as we can see the plots are overlaping, this shows us that there's no strong correlation between price and 
#### the table below apresents the mean between cut, note that there's not much difference between CUTS, this shows no strong correlation between factors (cut).
```{r}
by(diamonds$price,diamonds$cut, mean)
```
#### 7.5.2 two catergorical variables 

```{r}
library(tidyverse)
ggplot(data = diamonds) + geom_count(mapping = aes(x = cut, y = color))
```

#### 7.5.2.1 Exercises
##### How could you rescale the count dataset above to more clearly show the distribution of cut ##### within colour, or colour within cut?

```{r}
diamonds %>% group_by(color,cut) %>% summarise(n = n()) %>% arrange(-n)
```

```{r}
library(tidyverse)
diamonds %>% ggplot() + geom_bar(mapping = aes(x = reorder(cut, desc(cut)), fill = color))
```
```{r}
diamonds %>% 
  count(color, cut)
```
```{r}
ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(price, 2)))
```
#### 7.6

### relationship between faithful eruptions and waiting
```{r}
faithful %>% ggplot() + geom_point(mapping = aes(x = eruptions, y = waiting))
```

#### as we can see there's a relationship between eruptions and waiting, let's check how strong it is

```{r}
corrplot(cor(faithful), method  = 'number')
### or simply
cor(faithful$eruptions,faithful$waiting)
```
#### the relationship is considered strong r >= 0.7
#### Padrões apresentam uma das ferramentas mais úteis quando se tratando de data science pois eles revelam covariancia. Se você pensa a variancia como sendo um fenomeno que aumenta a incerteza, a covariancia é um fenomeno que a reduz. Se duas variáveis covariam, podemos usar o valor de uma variável para fazer melhores predições sobre os valores da segunda. Se a covariancia é devido a relação-causa podemos então usar o valor de uma variávle para controlar o valor da segunda; 

```{r}
mod = lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + geom_point(mapping = aes(x = carat, y = resid))

```
```{r}
diamonds2 %>% ggplot() + geom_boxplot(mapping = aes(x = cut, y = resid))

```



