---
title: "Chapter_5_R4DS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(nycflights13)
library(tidyverse)
rm(list=ls())
data = nycflights13::flights
##
```


##5.2.4 EXERCISES
####1. Find all flights that
      - Had an arrival delay of two or more hours
      - Flew to Houston (IAH or HOU)
      - Were operated by United, American, or Delta
      - Departed in summer (July, August, and September)
      - Arrived more than two hours late, but didn’t leave late
      - Were delayed by at least an hour, but made up over 30 minutes in flight
      - Departed between midnight and 6am (inclusive)
      
```{r}

prob1 = data[data$arr_delay >= 120,]
prob2 = data[data$dest == 'IAH' | data$dest == 'HOU',]

prob3 = data[data$carrier == 'AA' | data$carrier == 'UA' | data$carrier == 'DL',]
prob3b = data[data$carrier %in% c("AA","UA","DL"),]

prob4 = data[data$month %in% c(7,8,9),]
prob4b = data[data$month == 7 | data$month == 8 | data$month == 9,]

prob5 = data[data$arr_delay > 120 & data$dep_delay <= 0,]

prob6 = data[data$dep_delay > 60 & data$dep_time - data$arr_time >= 30,]
head(data)

prob7 = data[data$dep_time >= 2400 | data$dep_time <= 600,]

```

### 2 Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
  
```{r}
prob2= data[between(data$dep_time,2400,600),]
#no you cant use it.
```
### 3 How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
prob3 = colSums(is.na(data))
#8255, another variables are dep_delay,arr_time and so on
```
  
### 4 Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}
#these are operations that have the same output for every possible value.
```

##EXERCISES 5.3.1


### 1 How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
prob1 = data %>% filter(is.na(dep_time)) %>% arrange(dep_time)
```

### 2 Sort flights to find the most delayed flights. Find the flights that left earliest.


```{r}
prob2= data %>% arrange(min_rank(dep_delay))
```
### 3 Sort flights to find the fastest (highest speed) flights.

```{r}
prob3 = data %>% arrange(-air_time)
```

### 4 Which flights travelled the farthest? Which travelled the shortest?
```{r}
prob4 = data %>% group_by(flight) %>% select(distance, flight, everything()) %>% arrange(-distance)
prob4b = data %>%arrange(desc(air_time)) %>% select(air_time,everything())

```

### 1 Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights

```{r}
#many ways 
prob1 = data[,names(data) == c("dep_time", "dep_delay", "arr_time", "arr_delay")]
prob1b = data %>% select(dep_time,dep_delay,arr_time,arr_delay)
prob1c = data[,c(4,6,7,9)]
prob1d = data %>% select(contains(c('time', 'delay')))

```

### 2 What happens if you include the name of a variable multiple times in a select() call?
```{r}
data %>% select(arr_time, arr_time, arr_time)
#only works once.
```

### 3 What does the any_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
data %>% select(any_of(vars))
# it select all the variables the match the name in the variable.
```

### 4 Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
ex4 = data %>% select(dplyr::contains("TIME"))

#yes, i thought it was case sensitive, but it isnt.
```

### 5.5.2 Exercises

### 1 Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.


```{r}
ex1 = data %>% mutate (hour_deptime = dep_time %/% 100, minute_dep_time = dep_time %% 100) %>% select(hour_deptime, minute_dep_time,everything())

ex2 = data %>% mutate (hour_sch_deptime = sched_dep_time %/% 100, minute_sch_dep_time = sched_dep_time %% 100) %>% select(hour_sch_deptime, minute_sch_dep_time, everything())

```

### 2Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
ex1 =  data %>% select(air_time, arr_time, dep_time) %>%  mutate(arr_time-dep_time, arr_min = (arr_time %/% 100)*60 + (arr_time %% 100),dep_min = (dep_time %/% 100)*60 + (dep_time %% 100), trip_min = arr_min - dep_min)
ex1
```

#### 3 Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

dep delay = dep_time minus sched_dep_time as showed below however, we need to transform data first as shown below, in other words, transform it to correct hour/min using modulos and remainder operations and then subtract.
```{r}
ex3 = data %>% select(dep_time,sched_dep_time,dep_delay) %>% mutate(delay = dep_time - sched_dep_time, dep_timee_ajs = (dep_time %/% 100)*60 + (dep_time %% 100), sch_dep_time_min = (sched_dep_time %/% 100)*60 + (sched_dep_time %% 100), dep_delayy = dep_timee_ajs - sch_dep_time_min)
```

### 4 Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().
```{r}
data
data %>% mutate(rank = min_rank(dep_delay)) %>% arrange(-rank)
```
#Function min_rank is used along with mutate so in order to work it needs to be with mutate. It creates a new variable called rank, so its possible to rank this new variable and find the most delayed flight or vice versa.


### 5 What does 1:3 + 1:10 return? Why?
error because in order to sum these kind of object it must to be multiple. 
```{r}
1:3 + 1:10

#something that would work would be:
1:5 + 1:10

```
### 6 What trigonometric functions does R provide?

```{r}
    cos(x)
    sin(x)
    tan(x)
    acos(x)
    asin(x)
    atan(x)
    atan2(y, x)
    cospi(x)
    sinpi(x)
    tanpi(x)

```



### Chart to visualize the amount (count) of delayed flights using distance as parameter.
```{r}
delay = data %>% group_by(dest) %>% summarise(count = n(), dist = mean(distance, na.rm = T), delay = mean(arr_delay, na.rm = T)) %>% filter(count > 20, dest != 'HNL')

chart = ggplot(data = delay,aes(x = delay, y = dist)) + geom_point(aes(size = count))
chart

```
### 5.6.1 EXERCISES

### 1 Group flights by destination.

```{r}
names(data)
ex1 = data %>% group_by(dest) 
ex1
```
### 2 Summarise to compute distance, average delay, and number of flights.

```{r}
names(data)

data %>% group_by(dest) %>% summarise(soma_distancia = sum(distance, na.rm = T), 
                                      delay_medio = mean(arr_delay, na.rm = T),
                                      count_na = (sum(!is.na(dep_delay))),
                                      number_of_flights = n()) %>% arrange(-number_of_flights)

```


### 3 selecting only flights who havent got 'na' in dep_delay nor in arr_delay. 
```{r}
nao_cancelado <- data %>%
  filter(!is.na(air_time))
nao_cancelado
```


###5.6.7 EXERCISES

### Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

```{r}
nao_cancelado %>% 
  group_by(tailnum) %>% 
  mutate(
    count = n(),
    mediana_atraso_chegada = median(arr_delay),
    mediana_atraso_chegada = median(dep_delay)) %>% 
      filter(count > 5) %>% 
      arrange(mediana_atraso_chegada)
```

### 4 Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count())

```{r}
ex1 = data %>% group_by(dest) %>%  summarise(count = n())
ex2 = data %>% group_by(tailnum) %>% summarise(weight = sum(distance))
ex2
```


### 5.7.1 Exercises

### Which plane (tailnum) has the worst on-time record?

```{r}
data %>% group_by(tailnum) %>% summarise(media = round(mean(arr_delay),1)) %>% arrange(- media)
```
#### the worst one is N384HA which has got the most the biggest time, however if we compare the mean N844MH	is the tailnumber with the largest mean. 

### What time of day should you fly if you want to avoid delays as much as possible?

###teste



